Averaging subsampled pseudoinverses

### 2022-08-23 design/continuous_topml/appendix.tex
Commit original repos

\begin{proofof}{Lemma}{\ref{t:ring}} \
 First, we show that $\A+\u\v^\top$ is d.p.~for any fixed
 $\u,\v\in\R^d$. Below, we use a standard identity for the rank one
 update of a determinant:
 $\det(\A+\u\v^\top)=\det(\A)+\v^\top\!\adj(\A)\u$. It follows that
 for any $\Ic$ and $\Jc$ of the same size,
  \begin{align*}
\E\big[\!\det(\A_{\Ic,\Jc}\!+\u_{\Ic}\v_{\Jc}^\top)\big] &=
    \E\big[\!\det(\A_{\Ic,\Jc}) +
    \v_{\Jc}^\top\adj(\A_{\Ic,\Jc}) \u_{\Ic}\big]\\
    &\overset{(*)}{=}\det\!\big(\E[\A_{\Ic,\Jc}]\big) +
      \v_{\Jc}^\top\adj\!\big(\E[\A_{\Ic,\Jc}]\big) \u_{\Ic}\\
    &=\det\!\big(\E[\A_{\Ic,\Jc} \!+ \u_{\Ic}\v_{\Jc}^\top]\big),
  \end{align*}
  where $(*)$ used \eqref{eq:adj}, i.e., the fact that for d.p.~matrices, adjugate commutes
  with expectation. Crucially, through the definition of an adjugate
  this step implicitly relies on the assumption that all the square
  submatrices of $\A_{\Ic,\Jc}$ are also  determinant preserving.
  Iterating this, we get that $\A+\Z$ is d.p.~for any fixed
  $\Z$. We now show the same for $\A+\B$:
  \begin{align*}
\E\big[\!\det(\A_{\Ic,\Jc}\!+\B_{\Ic,\Jc})\big]
    &=
      \E\Big[\E\big[\!\det(\A_{\Ic,\Jc}\!+\B_{\Ic,\Jc})\mid\B\big]\Big]\\
    &\overset{(*)}{=}\E\Big[\!\det\!\big(\E[\A_{\Ic,\Jc}]\!+\B_{\Ic,\Jc}\big)\Big]\\
      &= \det\!\big(\E[\A_{\Ic,\Jc}\!+\B_{\Ic,\Jc}]\big),
  \end{align*}
  where $(*)$  uses the fact that after conditioning on $\B$ we can
  treat it as a fixed matrix. Next, we show that $\A\B$ is determinant preserving via the Cauchy-Binet formula:
  \begin{align*}
    \E\big[\!\det\!\big((\A\B)_{\Ic,\Jc}\big)\big]
    &= \E\big[\!\det(\A_{\Ic,*}\B_{*,\Jc})\big]\\
    &=\E\bigg[\sum_{S:\,|S|=|\Ic|}\!\!\det\!\big(\A_{\Ic,S}\big)
      \det\!\big(\B_{S,\Jc}\big)\bigg]\\
&=\!\!\sum_{S:\,|S|=|\Ic|}\!\!\det\!\big(\E[\A]_{\Ic,S}\big)
                                                \det\!\big(\E[\B]_{S,\Jc}\big)\\
    &=\det\!\big(\E[\A]_{\Ic,*}\, \E[\B]_{*,\Jc}\big)\\
      &= \det\!\big(\E[\A\B]_{\Ic,\Jc}\big),
  \end{align*}
  where recall that $\A_{\Ic,*}$ denotes the submatrix of $\A$
  consisting of its (entire) rows indexed by $\Ic$.
  \end{proofof}

To prove Lemma \ref{l:poisson}, we will use the following
lemma, many variants of which appeared in the literature
\cite[e.g.,][]{expected-generalized-variance}. We use the one given by
\cite{correcting-bias}.
\begin{lemma}[\citeauthor{correcting-bias}, \citeyear{correcting-bias}]\label{l:cb}
If the rows of random $k\times d$ matrices $\A,\B$
  are sampled as an i.i.d.~sequence of $k\geq d$ pairs of joint random vectors, then
\begin{align}
  k^d\,\E \big[\det(\A^\top\B)\big]
  &= \ktd\,\det\!\big(\E[\A^\top\B]\big).
     \end{align}
 \end{lemma}

@@

To prove Lemma \ref{l:normalization}, we use the following standard
determinantal formula which is used to derive the normalization
constant of a discrete determinantal point process.
\begin{lemma}[\citeauthor{dpp-ml}, \citeyear{dpp-ml}]\label{l:det-standard}
  For any $k\times d$ matrices $\A,\B$ we have
  \[\det(\I+\A\B^\top)=\sum_{S\subseteq[k]}\det(\A_{S,*}\B_{S,*}^\top).\]
\end{lemma}

@@

We first prove Lemma \ref{l:sqinv-all} by splitting the under- and
over-determined cases and start with proving the former.

@@

\begin{lemma}\label{l:sqinv-under}

@@

Applying the closed form expressions from Lemmas
\ref{l:proj} and \ref{l:sqinv-all}, we derive
the formula for the MSE and prove Theorem \ref{t:mse} (we defer the
proof of Lemma \ref{l:proj} to Appendix \ref{s:unbiased-proof}).
\begin{proofof}{Theorem}{\ref{t:mse}}
  First, assume that $n<d$, in which case we have
  $\gamma_n=\frac1{\lambda_n}$ and moreover
  \begin{align*}
    n &= \tr\big(\Sigmab_\mu(\Sigmab_\mu+\lambda_n\I)^{-1}\big)\\
      &=\tr\big((\Sigmab_\mu+\lambda_n\I-\lambda_n\I)(\Sigmab_\mu+\lambda_n\I)^{-1}\big)\\
    &=d - \lambda_n\tr\big((\Sigmab_\mu+\lambda_n\I)^{-1}\big),
    %\frac{\tr\big((\Sigmab_\mu+\lambda_n\I)^{-1}\big)}{d-n},
  \end{align*}
so we can write $\lambda_n$ as $(d-n)/\tr((\Sigmab_\mu+\lambda_n\I)^{-1})$.
  From this and Lemmas \ref{l:proj} and \ref{l:sqinv-under}, we
obtain the desired expression, where recall
  that $\alpha_n = \det\!\big(\Sigmab_\mu (\Sigmab_\mu+\frac1{\gamma_n})^{-1}\big)$:
  \begin{align*}
    \MSE{\Xb^\dagger\ybb} &= \sigma^2\,\gamma_n(1-\alpha_n) +
    \tfrac1{\gamma_n} \,\w^{*\top}(\Sigmab_\mu+\tfrac1{\gamma_n}\I)^{-1}\w^*
    \\
    &\overset{(a)}{=}\sigma^2\,\frac{1-\alpha_n}{\lambda_n} +
    \lambda_n\,\w^{*\top}(\Sigmab_\mu+\lambda_n\I)^{-1}\w^*\\
    &\overset{(b)}{=}\sigma^2\tr\big((\Sigmab_\mu+\lambda_n\I)^{-1}\big)\frac{1-\alpha_n}{d-n}
      +
      (d-n)\frac{\w^{*\top}(\Sigmab_\mu+\lambda_n\I)^{-1}\w^*}
      {\tr\big((\Sigmab_\mu+\lambda_n\I)^{-1}\big)}.
  \end{align*}
  While the expression given after $(a)$ is simpler than the one
after $(b)$, the latter better illustrates how the MSE depends on
the sample size $n$ and the dimension $d$.
  Now, assume that $n>d$. In this case, we have $\gamma_n=n-d$ and apply Lemma
  \ref{l:sqinv-over}:
  \begin{align*}
    \MSE{\Xb^\dagger\ybb}
    &= \sigma^2\,\tr(\Sigmab_\mu^{-1})\,
\frac{1-\ee^{-\gamma_n}}{\gamma_n}
=\sigma^2\,\tr(\Sigmab_\mu^{-1})
\,\frac{1-\beta_n}{n-d}.
  \end{align*}
The case of $n=d$ was shown in Theorem~2.12 of \cite{correcting-bias-journal}.
This concludes the proof.
\end{proofof}

@@

\section{Proof of Theorem \ref{t:asymptotic}}
\label{sec:proof-of-t-asymptotic}
Once again, we break down the proof into under- and
over-determined cases, starting with the former. Note that in this
case we require that the covariance be equal to identity.
\begin{lemma}\label{l:asymptotic-under}
  Let $\rho=n/d$, $\X\sim \Nc_{n,d}(\zero,\I_n \otimes \I_d)$ and $y_i=y(\x_i)$ under Assumption
  \ref{a:linear}. If $d>n+1$ then
  \begin{align*}
    &0\ \leq\ \frac{\MSE{\X^\dagger\y}-\Mc(\I,
\w^*,\sigma^2,n)}{\Mc(\I,\w^*,\sigma^2,n)}\ \leq\
\frac1d\cdot \frac{1}{1-\rho-\frac1d} + 3\rho^d.
  \end{align*}
\end{lemma}
  We first recall the standard decomposition of $\MSE{\X^\dagger \y}$:
    \MSE{\X^\dagger \y}
    &=\sigma^2 \E\big[\tr\big((\X^\top \X)^\dagger \big)\big]+
      \w^{*\top}\!\big( \I - \E[\X^\dagger \X]\big) \w^*.
  Since the rows of $\X$ are standard normal random variables,
  $\X\X^\top$ is an $n\times n$ Wishart random matrix with $d>n+1$ degrees
  of freedom. Using the formula for the mean of the Inverse-Wishart
  distribution, it follows that
\E\big[\tr((\X^\top\X)^\dagger)\big] = \E\big[\tr((\X\X^\top)^{-1})\big]= \frac{n}{d - n - 1}.
  % and therefore
  % \begin{align*}
  %   \MSE{\X^\dagger \y}
  %   &= \|\w\|_2^2 - \w^\top \E[\X^\dagger \X] \w
  %   + \frac{\sigma^2 n}{d - n - 1}
  % \end{align*}
  % Let $\X = \U \S \V^\top$ be its full singular value decomposition. Notice
  % \begin{align*}
  %   \X^\dagger \X
  %   &= \V \S^\dagger \U^\top \U \S \V^\top \\
  %   &= \V \S^\dagger \S \V^\top \\
  %   &= \V \begin{bmatrix}
  %     \I_n & 0 \\
  %     0 & 0
  %   \end{bmatrix} \V^\top \\
  %   &= \V_n \V_n^\top \\
  %   \w^\top \E[\X^\dagger \X] \w
  %   &= \E[\|\V_n^\top \w\|_2^2]
  % \end{align*}
  % where $\V_n$ denotes the $d \times n$ matrix with columns equal to the
  % $n$ eigenvectors of $\X^\top \X$ with non-zero eigenvalue.
  % But since $\Sigmab = \I$, by rotation invariance
  % $\V_n^\top$ is a change of basis onto a uniformly random
Note that $\X^\dagger\X$ is a uniformly random projection matrix so by
rotational symmetry it follows that
  \w^{*\top}\E\big[\X^\dagger\X]\w^* =
  \E\big[\|\X^\dagger\X\w^*\|^2\big] = \frac nd \,\|\w^*\|^2.
Putting this together we obtain that
    \MSE{\X^\dagger \y}
    &=\frac{\sigma^2 n}{d - n - 1}+ \|\w^*\|^2\,\frac{d - n}{d}.
  On the other hand, the surrogate MSE expression can be derived by
observing that for $\Sigmab=\I$ we have
$\tr((\Sigmab+\lambda_n\I)^{-1})=d/(1+\lambda_n)=n$ (see definition of
$\lambda_n$ in Theorem \ref{t:mse}):
    \Mc(\I, \w^*,\sigma^2,n) =
    \sigma^2n\cdot\frac{1-\alpha_n}{d-n} + \|\w^*\|^2\,\frac{d-n}{d}.
  Note that the second term is the same in both cases, even though
  this may not be true for non-isotropic Gaussians. We now compute the
  normalized difference between the
  expressions,
 \frac{\MSE{\X^\dagger\y}-\Mc(\I, \w^*,\sigma^2,n)}{\Mc(\I,
    \w^*,\sigma^2,n)}
    &= \frac{\sigma^2n\cdot(\frac1{d-n-1} -
    \frac{1-\alpha_n}{d-n})}{\sigma^2n\cdot\frac{1-\alpha_n}{d-n} +
    \|\w^*\|^2\,\frac{d-n}{d}}\\
&\leq \frac {d-n}{1-\alpha_n}
\Big(\frac1{d-n-1}-\frac{1-\alpha_n}{d-n}\Big)\\
    &=\frac{d-n}{d-n-1}\,\frac{1}{1-\alpha_n}- 1\\
    &=\frac1{d-n-1} + \frac{\alpha_n}{1-\alpha_n}\Big(1+\frac1{d-n-1}\Big).
Let $\rho=n/d$. Then $\frac1{d-n-1} = \frac1d\cdot
\frac1{1-\rho-\frac1d}$ and moreover
$\alpha_n = (\frac1{1+\lambda_n})^d=(\frac
nd)^d=\rho^d$. From the assumption that $d>n+1$, we
conclude that $\alpha_n \leq (\frac{d-2}{d})^d\leq \ee^{-2}$
so that $\frac{\alpha_n}{1-\alpha_n}(1+\frac1{d-n-1})\leq
  \frac{2 \alpha_n}{1-\ee^{-2}}\leq 3\rho^d$. This shows the right-hand-side inequality of the
  theorem. That fact that $\MSE{\X^\dagger\y}\geq\Mc(\I,
  \w^*,\sigma^2,n)$ follows easily.
\begin{lemma}\label{l:asymptotic-over}
  Let $\rho=n/d$, $\X\sim \Nc_{n,d}(\zero,\I_n \otimes \Sigmab)$ and $y_i=y(\x_i)$ under Assumption
  \ref{a:linear}. If $n>d+1$ then
    &0\ \leq\ \frac{\MSE{\X^\dagger\y}-\Mc(\Sigmab,
\w^*,\sigma^2,n)}{\Mc(\Sigmab,\w^*,\sigma^2,n)}\ \leq\
\frac1d\cdot \frac{1}{\rho-1-\frac1d} + 3(\ee^{1-\rho})^d.
The MSE for the over-determined Gaussian design can be derived by
using the formula for the mean of the Inverse-Wishart distribution:
  \MSE{\X^\dagger\y}=\sigma^2\tr\big(\E[(\X^\top\X)^{-1}]\big) =
  \frac{\sigma^2\tr(\Sigmab^{-1}) }{n-d-1}.
To compute the normalized difference we follow similar derivations as
in the proof of Lemma \ref{l:asymptotic-under}:
 \frac{\MSE{\X^\dagger\y}-\Mc(\Sigmab, \w^*,\sigma^2,n)}{\Mc(\Sigmab,
    \w^*,\sigma^2,n)}
    &=  \frac {n-d}{1-\beta_n}
      \Big(\frac1{n-d-1}-\frac{1-\beta_n}{n-d}\Big)\\
  &\leq\frac1d\cdot\frac1{\rho-1-\frac1d} + \frac{2\beta_n}{1-\beta_n}.
Recall that $\beta_n = \ee^{d-n} = (\ee^{1-\rho})^d$ and for $n-d\geq
2$ we have $\frac2{1-\beta_n}\leq 3$. The desired inequalities follow immediately.
Theorem \ref{t:asymptotic} now follows as a consequence of
Lemmas \ref{l:asymptotic-under} and \ref{l:asymptotic-over}.
\begin{proofof}{Theorem}{\ref{t:asymptotic}}
  Since $\frac1d\leq \frac13|1-\rho|$ and $\ee^{-|1-\rho|d}\leq
  \frac1{2d|1-\rho|}$, it follows that
    \frac1d\cdot \frac{1}{|1-\rho|-\frac1d} + 3(\ee^{-|1-\rho|})^d
    &\leq \frac1d\cdot \frac{1}{|1-\rho|-\frac13|1-\rho|} +
      \frac1d\cdot\frac{3}{2|1-\rho|}=\frac {c_\rho}{d}.
  The case of $n>d+1$ now follows from Lemma
  \ref{l:asymptotic-over}. Also, since $\rho\leq \ee^{\rho-1}$, for
$n<d-1$ we have $3\rho^d\leq 3(\ee^{-|1-\rho|})^{d}$, so the same bound follows
from Lemma \ref{l:asymptotic-under}.
\end{proofof}

\section{Empirical analysis of the asymptotic consistency conjectures}
\label{sec:asymp-conj-details}

>>>>

We use $\adj(\A)$ to denote the adjugate of $\A$, defined as follows: the
$(i,j)$th entry of $\adj(\A)$ is
$(-1)^{i+j}\det(\A_{[n]\backslash\{j\},[n]\backslash\{i\}})$.
We will use two useful identities related to the adjugate: (1)
$\adj(\A)=\det(\A)\A^{-1}$ for invertible $\A$, and (2)
$\det(\A+\u\v^\top)=\det(\A)+\v^\top\!\adj(\A)\u$
\citep[see Fact 2.14.2 in][]{matrix-mathematics}.

First, note that from the definition of an adjugate matrix it immediately follows that if $\A$ is
determinant preserving then adjugate commutes with expectation for this matrix:
\begin{align}
  \E\big[\big(\!\adj(\A)\big)_{i,j}\big] &=
  \E\big[(-1)^{i+j}\det(\A_{[d]\backslash\{j\},[d]\backslash\{i\}})\big]\nonumber
  \\
&=(-1)^{i+j}\det\!\big(\E[\A_{[d]\backslash\{j\},[d]\backslash\{i\}}]\big)
  \\
  &= \big(\!\adj(\E[\A])\big)_{i,j}.\label{eq:adj}
\end{align}
% The adjugate is useful in our analysis because it connects the
% determinant and the inverse via the formula
% $\adj(\A)=\det(\A)\A^{-1}$, which holds for any invertible $\A$.

\begin{proofof}{Lemma}{\ref{t:ring}} \
 First, we show that $\A+\u\v^\top$ is d.p.~for any fixed
 $\u,\v\in\R^d$. Below, we use the identity for a rank one
 update of a determinant:
 $\det(\A+\u\v^\top)=\det(\A)+\v^\top\!\adj(\A)\u$. It follows that
 for any $\Ic$ and $\Jc$ of the same size,
  \begin{align*}
\E\big[\!\det(\A_{\Ic,\Jc}\!+\u_{\Ic}\v_{\Jc}^\top)\big] &=
    \E\big[\!\det(\A_{\Ic,\Jc}) +
    \v_{\Jc}^\top\adj(\A_{\Ic,\Jc}) \u_{\Ic}\big]\\
    &\overset{(*)}{=}\det\!\big(\E[\A_{\Ic,\Jc}]\big) +
      \v_{\Jc}^\top\adj\!\big(\E[\A_{\Ic,\Jc}]\big) \u_{\Ic}\\
    &=\det\!\big(\E[\A_{\Ic,\Jc} \!+ \u_{\Ic}\v_{\Jc}^\top]\big),
  \end{align*}
  where $(*)$ used \eqref{eq:adj}, i.e., the fact that for d.p.~matrices, adjugate commutes
  with expectation. Crucially, through the definition of an adjugate
  this step implicitly relies on the assumption that all the square
  submatrices of $\A_{\Ic,\Jc}$ are also  determinant preserving.
  Iterating this, we get that $\A+\Z$ is d.p.~for any fixed
  $\Z$. We now show the same for $\A+\B$:
  \begin{align*}
\E\big[\!\det(\A_{\Ic,\Jc}\!+\B_{\Ic,\Jc})\big]
    &=
      \E\Big[\E\big[\!\det(\A_{\Ic,\Jc}\!+\B_{\Ic,\Jc})\mid\B\big]\Big]\\
    &\overset{(*)}{=}\E\Big[\!\det\!\big(\E[\A_{\Ic,\Jc}]\!+\B_{\Ic,\Jc}\big)\Big]\\
      &= \det\!\big(\E[\A_{\Ic,\Jc}\!+\B_{\Ic,\Jc}]\big),
  \end{align*}
  where $(*)$  uses the fact that after conditioning on $\B$ we can
  treat it as a fixed matrix. Next, we show that $\A\B$ is determinant preserving via the Cauchy-Binet formula:
  \begin{align*}
    \E\big[\!\det\!\big((\A\B)_{\Ic,\Jc}\big)\big]
    &= \E\big[\!\det(\A_{\Ic,*}\B_{*,\Jc})\big]\\
    &=\E\bigg[\sum_{S:\,|S|=|\Ic|}\!\!\det\!\big(\A_{\Ic,S}\big)
      \det\!\big(\B_{S,\Jc}\big)\bigg]\\
&=\!\!\sum_{S:\,|S|=|\Ic|}\!\!\det\!\big(\E[\A]_{\Ic,S}\big)
                                                \det\!\big(\E[\B]_{S,\Jc}\big)\\
    &=\det\!\big(\E[\A]_{\Ic,*}\, \E[\B]_{*,\Jc}\big)\\
      &= \det\!\big(\E[\A\B]_{\Ic,\Jc}\big),
  \end{align*}
  where recall that $\A_{\Ic,*}$ denotes the submatrix of $\A$
  consisting of its (entire) rows indexed by $\Ic$.
  \end{proofof}

To prove Lemma \ref{l:poisson}, we will use the following
lemma, many variants of which appeared in the literature
\cite[e.g.,][]{expected-generalized-variance}. We use the one given by
\cite{correcting-bias}.
\begin{lemma}[\cite{correcting-bias}]\label{l:cb}
If the rows of random $k\times d$ matrices $\A,\B$
  are sampled as an i.i.d.~sequence of $k\geq d$ pairs of joint random vectors, then
\begin{align}
  k^d\,\E \big[\det(\A^\top\B)\big]
  &= \ktd\,\det\!\big(\E[\A^\top\B]\big).
     \end{align}
 \end{lemma}

@@

To prove Lemma \ref{l:normalization}, we use the following standard
determinantal formula which is used to derive the normalization
constant of a discrete determinantal point process.
\begin{lemma}[\cite{dpp-ml}]\label{l:det-standard}
  For any $k\times d$ matrices $\A,\B$ we have
  \[\det(\I+\A\B^\top)=\sum_{S\subseteq[k]}\det(\A_{S,*}\B_{S,*}^\top).\]
\end{lemma}

@@

We first prove Lemma \ref{l:sqinv-all}. % by splitting the under- and
% over-determined cases and start with proving the former.

@@

\begin{lemma}[restated Lemma \ref{l:sqinv-all}]\label{l:sqinv-under}

@@

Applying the closed form expressions from Lemmas
\ref{l:proj}, \ref{l:sqinv-all} and \ref{l:sqinv-over}, we derive
the formula for the MSE and prove Theorem \ref{t:mse} (we defer the
proof of Lemma \ref{l:proj} to Appendix \ref{s:unbiased-proof}).
\begin{proofof}{Theorem}{\ref{t:mse}}
  First, assume that $n<d$, in which case we have
  $\gamma_n=\frac1{\lambda_n}$ and moreover
  \begin{align*}
    n &= \tr\big(\Sigmab_\mu(\Sigmab_\mu+\lambda_n\I)^{-1}\big)\\
      &=\tr\big((\Sigmab_\mu+\lambda_n\I-\lambda_n\I)(\Sigmab_\mu+\lambda_n\I)^{-1}\big)\\
    &=d - \lambda_n\tr\big((\Sigmab_\mu+\lambda_n\I)^{-1}\big),
    %\frac{\tr\big((\Sigmab_\mu+\lambda_n\I)^{-1}\big)}{d-n},
  \end{align*}
so we can write $\lambda_n$ as $(d-n)/\tr((\Sigmab_\mu+\lambda_n\I)^{-1})$.
  From this and Lemmas \ref{l:proj} and \ref{l:sqinv-under}, we
obtain the desired expression, where recall
  that $\alpha_n = \det\!\big(\Sigmab_\mu (\Sigmab_\mu+\frac1{\gamma_n})^{-1}\big)$:
  \begin{align*}
    \MSE{\Xb^\dagger\ybb} &= \sigma^2\,\gamma_n(1-\alpha_n) +
    \tfrac1{\gamma_n} \,\w^{*\top}(\Sigmab_\mu+\tfrac1{\gamma_n}\I)^{-1}\w^*
    \\
    &\overset{(a)}{=}\sigma^2\,\frac{1-\alpha_n}{\lambda_n} +
    \lambda_n\,\w^{*\top}(\Sigmab_\mu+\lambda_n\I)^{-1}\w^*\\
    &\overset{(b)}{=}\sigma^2\tr\big((\Sigmab_\mu+\lambda_n\I)^{-1}\big)\frac{1-\alpha_n}{d-n}
      +
      (d-n)\frac{\w^{*\top}(\Sigmab_\mu+\lambda_n\I)^{-1}\w^*}
      {\tr\big((\Sigmab_\mu+\lambda_n\I)^{-1}\big)}.
  \end{align*}
  While the expression given after $(a)$ is simpler than the one
after $(b)$, the latter better illustrates how the MSE depends on
the sample size $n$ and the dimension $d$.
  Now, assume that $n>d$. In this case, we have $\gamma_n=n-d$ and apply Lemma
  \ref{l:sqinv-over}:
  \begin{align*}
    \MSE{\Xb^\dagger\ybb}
    &= \sigma^2\,\tr(\Sigmab_\mu^{-1})\,
\frac{1-\ee^{-\gamma_n}}{\gamma_n}
=\sigma^2\,\tr(\Sigmab_\mu^{-1})
\,\frac{1-\beta_n}{n-d}.
  \end{align*}
The case of $n=d$ was shown in Theorem~2.12 of \cite{correcting-bias-journal}.
This concludes the proof.
\end{proofof}

@@

\section{Proof of Theorem \ref{t:asymptotic}}
\label{sec:proof-of-t-asymptotic}

The proof of Theorem \ref{t:asymptotic} follows the standard decomposition of
MSE in Equation~\ref{eq:mse-derivation}, and in the process,
establishes consistency of the variance and bias terms
independently. To this end, we introduce the following two useful
lemmas that capture the limiting behavior of the 
variance and bias terms, respectively.

\begin{lemma}\label{c:wishart}
  Under the setting of Theorem~\ref{t:asymptotic}, we have, as $n,d \to \infty$
with $n/d \to \bar c \in (0, \infty) \setminus \{ 1 \}$ that
\begin{align}
%\bigg|\frac{\E\big[\tr((\X^\top \X)^\dagger)\big]}{\Vc(\Sigmab,n)} -1\bigg| \to 0 \qquad\text{for}\quad\Vc(\Sigmab,n)= \frac{1-\alpha_n}{\lambda_n}
\begin{cases}
  \E\big[\tr((\X^\top \X)^\dagger)\big] - (1-\alpha_n) \lambda_n^{-1} \to 0, & \text{for }\bar c < 1, \\
  \E\big[\tr((\X^\top \X)^\dagger)\big] - \frac{1-\beta_n}{n - d} \cdot \tr \Sigmab^{-1} \to 0, & \text{for }\bar c > 1
\end{cases}
\label{eq:wishart}
\end{align}
where $\lambda_n\geq 0$ is the unique solution to
$n=\tr(\Sigmab(\Sigmab+\lambda_n\I)^{-1})$,
$\alpha_n=\det(\Sigmab(\Sigmab+\lambda_n\I)^{-1})$,
and $\beta_n = e^{d-n}$.
\end{lemma}

\noindent

The second term in the MSE derivation \eqref{eq:mse-derivation}, $\E[\I -
\X^\dag \X]$, involves the expectation of a projection onto the orthogonal
complement of a sub-Gaussian general position sample $\X$.
This term is zero when $n > d$, and for $n < d$
we prove in \cref{sec:proof-of-c-projection} that the surrogate design's
bias $\mathcal{B}(\Sigmab, n)$ provides an asymptotically consistent approximation
to all of the eigenvectors and eigenvalues:

\begin{lemma}\label{c:projection}
Under the setting of Theorem~\ref{t:asymptotic}, for $\w \in \R^d$ of bounded Euclidean norm (i.e., $\| \w \| \le C'$ for all $d$), we have, as $n,d \to \infty$ with $n/d \to \bar c \in (0,1)$ that
\begin{align}
%\sup_{\w\in\R^d\backslash\{\zero\}}\bigg|\frac{\w^\top\E[\I-\X^\dagger\X]\w}{\w^\top \Bc(\Sigmab,n)\w} - 1\bigg| \to 0 \qquad\text{for}\quad\Bc(\Sigmab,n) = \lambda_n(\Sigmab+\lambda_n\I)^{-1}
  \w^\top\E[\I-\X^\dagger\X]\w - \lambda_n \w^\top (\Sigmab + \lambda_n \I)^{-1}\w \to 0
  \label{eq:projection}
\end{align}
while $\I - \X^\dagger \X = 0$ for $\bar c > 1$.
\end{lemma}

\subsection{Proof of \cref{c:wishart}}
\label{sec:proof-of-c-wishart}

\subsubsection{The $\bar c \in (0,1)$ case}

For $n < d$, we first establish (1) $\liminf_n \lambda_n > 0$ and (2) $\alpha_n \to 0$.
To prove (1), by hypothesis $\Sigmab \succeq c \I$ for all $d$. Since $\frac{n}{d} < 1$,
we have (by definition of $\lambda_n$) for some $\delta > 0$
\begin{align*}
  1 - \delta
  > \frac{n}{d}
  = \frac{1}{d} \tr(\Sigmab(\Sigmab + \lambda_n \I)^{-1})
  > \frac{c}{c + \lambda_n}
\end{align*}
Rearranging, we have $\lambda_n > \frac{\delta c}{1 - \delta} > 0$.
For (2), let $(\tau_i)_{i \in [d]}$ denote the eigenvalues of $\Sigmab$.
Since $1 - x \leq e^{-x}$ and $C \I \succeq \Sigmab \succeq c \I$ for all $d$,
\begin{align*}
  \alpha_n
  = \prod_{i=1}^d \frac{\tau_i}{\tau_i + \lambda_n}
  \leq \left(\frac{C}{C + \lambda_n}\right)^d
  = \left( 1 - \frac{\lambda_n}{C + \lambda_n}\right)^d
  \leq \exp\left(-d \frac{\lambda_n}{C + \lambda_n}\right)
\end{align*}
and since $\lambda_n > 0$ eventually as $d \to \infty$ we have
$\alpha_n \to 0$ so that $(1-\alpha_n) \lambda_n^{-1} - \lambda_n^{-1} \to 0$.

% (1) and (2) together imply $\mathcal{V}(\Sigmab, n) \to \lambda_n^{-1}$, and
% by Slutsky's theorem
% \begin{align*}
%   \lim_{n,d} \left\lvert
%   \frac{\E[\tr((\X^\top \X)^\dag)] - \mathcal{V}(\Sigmab, n)}{\mathcal{V}(\Sigmab, n)}
%   \right\rvert
%   = \lim_{n,d} \left\lvert
%   \frac{\E[\tr((\X^\top \X)^\dag)] - \lambda_n^{-1}}{\lambda_n^{-1}}
%   \right\rvert
% \end{align*}
% Since $\liminf_n \lambda_n > 0$ was shown in (1),
As a consequence of (2) and Slutsky's theorem, it suffices to show
$\tr(\X^\top \X)^\dag - \lambda_n^{-1} \overset{d}{\to} 0$ as $n,d \to \infty$.
To do this, we consider the limiting behavior of $\tr
(\X^\top \X)^\dagger/n = \tr (\X \X^\top)^\dagger/n$ as $n/d \to \bar c \in
(0,1)$, for $\X = \Z \Sigmab^{\frac12}$ with $\Z \in \mathbb R^{n \times d}$
having i.i.d.~zero mean, unit variance sub-Gaussian entries, i.e., the behavior
of
\begin{equation}\label{eq:limit1}
  \lim_{n,d \to \infty} \lim_{z \to 0^+} \frac1n \tr \left( \frac1n \X \X^\top + z \I_n \right)^{-1}
\end{equation}
by definition of the pseudo-inverse.

The proof comes in three steps: (i) for fixed $z > 0$, consider the limiting behavior of $ \delta(z) \equiv \tr (\X \X^\top/n + z \I_n)^{-1}/n$ as $n,d \to \infty$ and state
\begin{equation}\label{eq:limit2}
  \lim_{n,d \to \infty} \delta(z) - m(z) \to 0
\end{equation}
almost surely for some $m(z)$ to be defined; (ii) show that both $\delta(z)$ and its derivate $\delta'(z)$ are uniformly bounded (by some quantity independent of $z>0$) so that by Arzela-Ascoli theorem, $\delta(z)$ converges uniformly to its limit and we are allowed to take $z \to 0^+$ in \eqref{eq:limit2} and state
\begin{equation}\label{eq:limit3}
  \lim_{z \to 0^+} \lim_{n,d \to \infty} \delta(z) - \lim_{z \to 0^+} m(z) \to 0
\end{equation}
almost surely, given that the limit $\lim_{z \to 0^+} m(z) \equiv m(0)$ exists and eventually (iii) exchange the two limits in \eqref{eq:limit3} with Moore-Osgood theorem, to reach
\[
  \lim_{n,d \to \infty} \lim_{z \to 0^+} \frac1n \tr \left(\frac1n \X \X^\top + z \I_n \right)^{-1} - m(0) \to 0.
\]

Step (i) follows from \cite{silverstein1995empirical} that, we have, for $z > 0$ that
\[
  \delta(z) \equiv \frac1n \tr \left( \frac1n \X \X^\top  + z \I_n \right)^{-1}  - m(z) \to 0
\]
almost surely as $n,d \to \infty$, for $m(z)$ the unique positive solution to
\begin{equation}\label{eq:def-m}
  m(z) = \left( z + \frac1n \tr \Sigmab (\I + m(z) \Sigmab)^{-1} \right)^{-1}.
\end{equation}

For the above step (ii), we use the assumption $\Sigmab \succeq c \I \succ 0$
for all $d$ large, so that with $\X = \Z \Sigmab^{\frac12}$, we have for large enough $n,d$ that
\[
  \lambda_{\min} (\X \X^\top/n) \ge \lambda_{\min} (\Z \Z^\top/n) \lambda_{\min} (\Sigmab) \ge \frac{c}2 (\sqrt{\bar c} - 1 )^2
\]
almost surely, where we used Bai-Yin theorem \cite{bai1993limit}, which states that the minimum eigenvalue of $\Z \Z^\top/n$ is almost surely larger than $( \sqrt{\bar c} - 1 )^2/2$ for $n<d$ sufficiently large. Note that here the case $\bar c = 1$ is excluded.

Observe that
\[
  |\delta(z)| = \left| \frac1n \tr \left( \frac1n \X \X^\top  + z \I_n \right)^{-1} \right| \le \frac1{ \lambda_{\min} (\X \X^\top/n) }
\]
and similarly for its derivative, so that we are allowed to take the $z \to 0^+$ limit. Note that the existence of the $\lim_{z \to 0^+} m(z)$ for $m(z)$ defined in \eqref{eq:def-m} is well known, see for example \cite{ledoit2011eigenvectors}. Then, by Moore-Osgood theorem we finish step (iii) and by concluding that
\[
  \tr (\X^\top \X)^\dagger - m(0) \to 0
\]
for $m(0) = \lambda_n^{-1}$ the unique solution to $\lambda_n^{-1} = \left( \frac1n \tr \Sigmab (\I + \lambda_n^{-1} \Sigmab)^{-1} \right)^{-1}$, or equivalently, to
\[
  n = \tr \Sigmab (\Sigmab + \lambda_n \I)^{-1}
\]
as desired.

\subsubsection{The $\bar c \in (1, \infty)$ case}

First note that as $n, d \to \infty$ with $n > d$, we have $\beta_n = e^{d-n} \to 0$ and it
it suffices to show
\[
  \tr (\X^\top \X)^\dagger - \frac1{n-d} \tr \Sigmab^{-1} \to 0
\]
almost surely to conclude the proof.
